---
title: "MiniProject3"
author: "Tyson Hyland, Makena Mecham"
format: html
---

```{r}
library(tidymodels)
library(tidyverse)
library(workflows)
library(tune)
library(mlbench)
```

## a. Split the data into `training` and `test` sets

```{r}
heart_data <- read_csv('./data/heart_failure_clinical_records_dataset.csv')
heart_data$DEATH_EVENT <- as.factor(heart_data$DEATH_EVENT) # Convert the response to a factor to work in the rf model later
# any(is.na(heart_data)) # A check for NA values reveals there are none in the data
set.seed(123) # Maintain a consistent seed
heart_split <- initial_split(heart_data, strata = DEATH_EVENT, prop = 4/5) # Split the data into training and testing sets
train <- training(heart_split) # Training data
test <- testing(heart_split) # Testing data
heart_cv <- vfold_cv(train) # Crossvalidation fold
```

## b. Build a recipe using the recipe() function and the step\_\*() functions

```{r}
heart_recipe <- recipe(DEATH_EVENT ~ ., data=heart_data) %>%
  step_normalize(all_numeric()) %>%
  step_impute_knn(all_predictors())

# heart_recipe
```

```{r}
heart_process <- heart_recipe %>%
  prep(train) %>%
  juice()
# heart_process
```

## c. Specify the model using the rand_forest() and set_engine functions.

```{r}
rf_model <- 
  # specify that the model is a random forest
  rand_forest() %>%
  # specify that the `mtry` parameter needs to be tuned
  set_args(mtry = tune(),trees = tune(),) %>%
  # select the engine/package that underlies the model
  set_engine("ranger", importance = "impurity") %>%
  # choose either the continuous regression or binary classification mode
  set_mode("classification") 
```

## d. Create a workflow and add the recipe.

```{r}
rf_workflow <- workflow() %>% 
  add_recipe(heart_recipe) %>% 
  add_model(rf_model)
# rf_workflow
```

## e. Tune the model using tune_grid() to find the optimal value of the tuning parameter mtry (the number of randomly selected predictors at each split) using 5-fold CV, where mtry varies over 1, 5, 9, ... , 𝑝. Choose the best value using select_best().

```{r}
doParallel::registerDoParallel()
set.seed(123)

p <- ncol(heart_data) - 1

rf_grid <- expand_grid(mtry = seq(1, p, by = 4), trees = c(100, 500)) # Passing values 
# extract results
rf_tune_results <- rf_workflow %>%
  tune_grid(resamples = heart_cv, #CV object
            grid = rf_grid, # grid of values to try
            metrics = metric_set(accuracy, roc_auc) # metrics we care about
            )

# rf_tune_results %>% collect_metrics()

param_final <- rf_tune_results %>% 
  select_best(metric='roc_auc')
# param_final
```

## f. Evaluate the performance of your final model using the finalize_workflow function and add_model on the test set and report the optimal value of mtry, the confusion matrix and the misclassification error rate. Use set.seed(123) for reproducibility of your results.

```{r}
# Finalize the workflow
rf_workflow <- rf_workflow %>%
  finalize_workflow(param_final)
```

```{r}
# Fit the workflow to a model and test/train on the split data
rf_fit <- rf_workflow %>%
  # fit on the training set and evaluate on test set
  last_fit(heart_split)

#TODO extract the value of mtry() from this model

test_performance <- rf_fit %>% collect_metrics()
test_performance
```

```{r}
ranger_obj <- extract_fit_parsnip(rf_fit)$fit
ranger_obj
```

```{r}
# Convert variable importance to a data frame
importance_df <- as.data.frame(ranger_obj$variable.importance) %>%
  tibble::rownames_to_column("Variable") %>%  # Add variable names as a column
  dplyr::rename(Importance = `ranger_obj$variable.importance`) %>%  # Rename the importance column
  dplyr::arrange(desc(Importance))  # Sort by importance, if desired

# Print the formatted data frame
print(importance_df)
```

```{r}
# generate predictions from the test set
test_predictions <- rf_fit %>% collect_predictions()
test_predictions
```

```{r}
# generate a confusion matrix
test_predictions %>% 
  conf_mat(truth = DEATH_EVENT, estimate = .pred_class)
```

Here, the mtry value in our final model is 1. The accuracy is 83.6%, and the roc accuracy is 94.6%. The misclassification error rate is 10/61 = 16.4%.

## g. Plot the ROC curve.

```{r}

```

## h. Plot the important variables

```{r}
# Convert variable importance to a data frame
importance_df <- as.data.frame(ranger_obj$variable.importance) %>%
  tibble::rownames_to_column("Variable") %>%  # Add variable names as a column
  dplyr::rename(Importance = `ranger_obj$variable.importance`) %>%  # Rename the importance column
  dplyr::arrange(desc(Importance))  # Sort by importance, if desired

# Print the formatted data frame
print(importance_df)
```

```{r}
# Assuming your dataframe is named importance_df
ggplot(importance_df, aes(x = reorder(Variable, Importance), y = Importance)) +
  geom_col(fill = "steelblue") +  # Bar plot with custom color
  coord_flip() +  # Flip coordinates for easier readability
  labs(
    title = "Variable Importance",
    x = "Variable",
    y = "Importance"
  ) +
  theme_minimal()
```
